Detection of slow slip events using wavelet analysis of GNSS recordings

### 2022-06-29 paper/paper_bssa_annotated.tex
Ready to submit

\begin{itemize}
\item We use a wavelet-based signal processing method to detect transients in GNSS data, such as slow slip events.
\item There is a good correlation between detections of slow slip using GNSS data and using tremor data.
\item The method could be applied in regions where no tremor is detected in conjunction with slow slip events.
\end{itemize}

@@

In many places, tectonic tremor is observed in relation to slow slip and can be used as a proxy to study slow slip events of moderate magnitude where surface deformation is hidden in Global Navigation Satellite System (GNSS) noise. However, in subduction zones where no clear relationship between tremor and slow slip occurrence is observed, these methods cannot be applied, and we need other methods to be able to better detect and quantify slow slip. Wavelets methods such as the Discrete Wavelet Transform (DWT) and the Maximal Overlap Discrete Wavelet Transform (MODWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted differences of a time series vary from one period to the next. In this paper, we use wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia. We use detrended GNSS data, apply the MODWT transform and stack the wavelet details over several nearby GNSS stations. As an independent check on the timing of slow slip events, we also compute the cumulative number of tremor in the vicinity of the GNSS stations, detrend this signal, and apply the MODWT transform. In both time series, we can then see simultaneous waveforms whose timing corresponds to the timing of slow slip events. We assume that there is a slow slip event whenever there is a positive peak followed by a negative peak in the wavelet signal. We verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data for northern Cascadia. The wavelet-based detection method detects well events of magnitude higher than 6 as determined by independent event catalogs (e.g. ~\citep{MIC_2019}).

@@

Slow slip events are a new feature discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks ~\citep{VER_2010,SCH_2010,JIA_2012,WAL_2012}. As with ordinary earthquakes, slow slip events represent slip on a fault, for instance the plate boundary between a tectonic plate subducting under another tectonic plate. However, they take a much longer time (several days to several years) to happen relative to ordinary earthquakes. They have a relatively short recurrence time (months to years) compared to the recurrence time of regular earthquakes (up to several hundreds of years), allowing scientists to observe and study many complete event cycles, which is typically not possible to explore with traditional earthquake catalogs ~\citep{BER_2011}. A slow slip event on the plate boundary is inferred to happen when there is a reversal of the direction of motion at GNSS stations, compared to the secular interseismic motion. Slow slip events have been observed in many places ~\citep{BER_2011,AUD_2016}, such as Cascadia ~\citep{BAR_2020}, Nankai ~\citep{NIS_2013}, Alaska ~\citep{LI_2016}, Costa Rica ~\citep{JIA_2012}, Mexico ~\citep{RAD_2012}, and New Zealand ~\citep{WAL_2020}. \\

In many places, tectonic tremor is also observed in relation to slow slip, but it is more abundant in some places ~\citep{HAL_2018}. Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases. Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs) ~\citep{SHE_2007_nature}, which are small magnitude earthquakes (M $\sim$ 1) for which frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz). In subduction zones such as Nankai and Cascadia, tectonic tremor observations are spatially and temporally correlated with slow slip observations ~\citep{OBA_2002,ROG_2003}. Due to this correlation, these paired phenomena have been called Episodic Tremor and Slip (ETS). However, this is not always the case. For instance, in northern New Zealand, tremor is more challenging to detect, and seems to be located downdip of the slow slip on the plate boundary ~\citep{TOD_2016}. In Alaska, the tremor zone only partially overlaps the long-term slow slip zone and there does not appear to be any temporal correlation between tremor and slow slip occurrence ~\citep{WEC_2016}. \\

In Cascadia, there are robust signals in both GNSS and tremor ~\citep{HAW_2013}. This is also the case in Nankai ~\citep{HIR_2008}, where tiltmeters are used instead of GNSS. It is thus possible to use tremor as a proxy to observe slow slip events that are not directly observed in the GNSS data. For instance, ~\citet{AGU_2009} studied 23 ETS events in Cascadia with more than 50 hours of tectonic tremor. For all these events, they computed both the GPS-estimated moment release and the cumulative number of hours of tectonic tremor recorded. They observed a linear relationship between moment release and number of hours of tremor for ETS events of moment magnitude 6.3 to 6.8. Based on this linear relationship, it is possible to infer the existence of smaller slow slip events of magnitude 5-6 occurring simultaneously with smaller tremor bursts of duration 1 to 50 hours occurring in between the big ETS events, and for which there is no detectable signal in the GPS data. \\

@@

Finally, ~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine relative vertical displacements associated with ETS events between 1996 and 2011. Their main idea is that the tidal level measured at a given gauge is the sum of a noise component at multiple timescales (tides, ocean and atmospheric noise) and an uplift signal due to the ETS events. The noise component is assumed to be coherent between all tidal gauges, while the tectonic uplift signal is different provided that the gauges are far enough from each other. By stacking the tidal records after removing tides, the uplift signals cancel each other while the noise signal is amplified. By stacking the details of the DWT decomposition, instead of stacking the raw tidal record, each of the components of the noise at different time scales is retrieved and can then be removed from the raw records to obtain the uplift signal. The authors were then able to clearly see a difference in uplift between the two tidal gauges at Port Angeles and Port Townsend. \\

In our study, we use a similar approach to previous studies with a different reasoning. We only stack signals at nearby GPS stations, assuming that the longitudinal displacement due to the ETS events will then be the same at each of the GPS stations considered. We suppose that some of the noise component is different at each GPS station and will be eliminated by the stacking. Finally, we assume that the noise and the longitudinal displacement due to the ETS events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the ETS events. We use wavelet methods to analyze GPS and tremor recordings of slow slip events in Cascadia. Our objective is to verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data. We thus want to demonstrate that the wavelet-based detection method can be applied to detect slow slip events that may currently be obscured using standard methods. \\

@@

@@

The DWT and MODWT methods must be used on a continuous time series, without gaps in the recordings. To deal with the gaps in the GNSS recordings, we simply replace the missing values by interpolation. The value for the first day for which data are missing is equal to the mean of the five days before the gap. The value for the last day for which data are missing is equal to the mean of the five days after the gap. The remaining missing values are computed by doing a linear interpolation of the first and the last values and adding a Gaussian noise component with mean zero and standard deviation equal to the standard deviation of the whole time series. The straight line starts at and ends at . We verify how the wavelet details may be affected by looking at a GPS time series without missing values and compared the wavelet details with and without removing some data points. Station PGC5 recorded continuous 1390 days between 2009 and 2013 without any missing values. We first computed the wavelet details without missing values. Then, we removed ten neighboring values, replaced them using the method described above (linear interpolation plus Gaussian noise), and computed the wavelet details with the replaced values. Figure 2 shows a comparison of the two wavelet details for two different locations of the missing values. We can see that there are visible differences in the time series itself, and in the details at the smallest levels of the wavelet decomposition. However, the differences between the wavelet details with and without missing values get smaller and smaller with increasing levels of details, and are barely visible for the levels that are most relevant (levels 6 and above). We thus conclude that we can easily replace the missing values in the GNSS time series without introducing false detections of slow slip events. \\

We then applied the wavelet filtering to real GPS data. Figure 3 shows the longitudinal displacement for GPS station PGC5, located in southern Vancouver Island, the details of the wavelet decomposition for levels 1 to 8, and the smooth. In the data, we can see a sharp drop in displacement whenever there is a documented slow slip event. For levels 5 to 8, which correspond to time scales 16, 32, 64 and 128 days, we can see in the details a positive peak followed by a negative peak whenever there is a drop in displacement in the data. We thus verify that the wavelet method can detect steps in the time series associated with slow slip events. \\

To increase the signal-to-noise ratio and better detect slow slip events, we stack the signal from several neighboring GPS stations. We choose to focus on GPS stations located close enough to the tremor zone to get a sufficiently high amplitude of the slow slip signal. We choose 16 points along the 40 km depth contour of the plate boundary (model from ~\citet{PRE_2003}) with spacing equal 0.1 degree in latitude (red triangles on Figure 4). Then we took all the GPS stations located in a 50 km radius for a given point, compute the wavelet details for the longitudinal displacement of each station, and stack each detail over the GPS stations. We thus have a stacked detail for each level 1 to 10 of the wavelet decomposition. \\

To assess the success of the wavelet decomposition for detecting slow slip events in GPS time series, we validate the approach by comparing to an independent proxy for ETS events. We took all the tremor epicenters located within a 50 km radius centered on one of the 16 locations marked by red triangles on Figure 3. Then we computed the cumulative number of tremor within this circle. Finally, we removed a linear trend from the cumulative tremor count, and applied the wavelet transform. Figure 5 shows an example of the wavelet decomposition for the third northernmost location on Figure 4 (which is closest to GPS station PGC5). Contrary to what happens for the GPS data, we see a sharp increase in the time series whenever there is a tremor episode, which translates into a negative peak followed by a positive peak in the wavelet details.

\section{Results}

We stacked the 8th level detail of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. The result is shown in the top panel of Figure 6, where each line represents one of the locations along strike. To better highlight the peaks in the wavelet details, we highlighted in red the time intervals where the amplitude of the stacked detail is higher than a threshold, and in blue the  time intervals where the amplitude of the stacked detail is lower than minus the threshold. To compare the GPS signal with the tremor signal, we plotted the 8th level detail of the wavelet decomposition of the tremor count on the bottom panel of Figure 6. We multiplied by -1 the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks. In the tremor catalog from the PNSN, there are 17 tremor events with more than 150 hours of tremor recorded. The events are summarized in Table 1. The time of the event is the start date plus half the duration of the event. \\

@@

We can then evaluate the quality of the detections obtained with our method by plotting a receiver operating characteristic curve (ROC curve). The ROC curve is widely use for binary classification problems in statistics and machine learning. We calculate an ROC value by varying the values of the threshold (here the two thresholds used to convert the GPS and the tremor time series into binary time series), computing the corresponding values of the true positive rate and the false positive rate (equal to 1 minus the specificity), and plotting the true positive rate as a function of the false positive rate. If the classification was made randomly, all the points would fall on the first diagonal. If the classifier was perfect, the corresponding point would fall on the top left corner of the graph with true positive rate equal to 1 and false positive rate equal to 0. The bigger the area under the curve, the better the classification method is. \\

As the slow slip events are better seen on levels 6, 7 and 8 of the wavelet decomposition, we first add the wavelet details corresponding to levels 6 to 8, and transform the resulting time series into a binary time series. We apply this transform to both the GPS and the tremor time series with varying thresholds. We then plot the ROC curve on Figure 9, each dot representing a different threshold. The corresponding sums of the wavelet details for the GPS data and the tremor data are shown on Figure 10. We can see that there is a trade-off between sensitivity and specificity as we vary the threshold. If we decrease the false positive rate, we also decrease the number of true events detected. If we increase the number of true events detected, we also increase the false positive rate. In Figure 10, we have chosen thresholds for the GPS time series and the tremor time series such that the specificity is higher than 0.75, and the sensitivity is the highest possible, that is we have chosen the thresholds corresponding to the dot that is farthest from the diagonal, which is random. \\

In addition to the magnitude 6 events discussed above,  ~\citet{MIC_2019} have also identified several magnitude 5 events using a variational Bayesian Independent Component Analysis (vbICA) decomposition of the signal. As we expect smaller magnitude events to be more visible at smaller time scales of the wavelet decomposition (level 5), we verify for all these events whether a signal can be seen at the same time as the time given in their catalog. Most of these magnitude 5 events are also sub-events of bigger magnitude 6 events. Table 2 summarizes for each event its timing, its number and its magnitude as indicated in the catalog from ~\citet{MIC_2019}, and whether it is part of a bigger magnitude 6 event. Figure 11 shows the 5th level detail wavelet decomposition of the GPS data. Red lines show the timing of the big ETS events from Table 1, and blue lines show the timing of the small slow slip events from Table 2. \\

All 14 events that are sub-events of a bigger event are visible at level 5. However, this may be because the bigger events are clearly seen at levels 6 to 8, and also at smaller time scales. The one small event that is not part of a bigger event (Winter 2009) is visible at level 5 of the wavelet decomposition. However, some other events that are not in the catalog of ~\citet{MIC_2019}'s catalog are also visible in late 2007, early 2010, early 2012, and late 2016. Therefore, it is difficult to differentiate between a true detection and a false detection, and to conclude whether the method can indeed detect events of magnitude 5. \\

In Figure 9, we see four smaller events that are not in the catalog of ~\citet{MIC_2019}: at about 2007.5, there is a negative peak followed by a positive peak (that is an event in the opposite direction of what would be expected from slow slip), at about 2010.2, 2012.2 and 2020.2, there are positive peaks followed by negative peaks for all the sixteen locations studied in this paper. Looking back at the original GPS data, there is a small increase in the displacement in the eastern direction that lasts about one or two months at about 2007.5. However, the direction of the displacement does not correspond to a slow slip event, and another cause should be found to explain this signal. There is a decrease in displacement that lasts several months at about 2010.2. This transient may correspond to a long duration slow slip event. There is a small decrease in displacement at about 2012.2. Its amplitude is small but the duration and direction correspond to a slow slip event, so this transient could be a very small slow slip event. Finally, there is also a small decrease in displacement at about 2020.2 that is difficult to interpret.

>>>>

\begin{itemize}
\item We use a wavelet-based signal processing method to detect transients in GNSS data, such as slow slip events.
\item In Northern Cascadia, there is good \textbf{agreement} between detections of slow slip using GNSS data and using tremor data.
\item The method \textbf{can} be applied \textbf{as a starting point to identify slow slip without prior information. New Zealand is shown as on example.}
\end{itemize}

@@

In many places, tectonic tremor is observed in relation to slow slip and can be used as a proxy to study slow slip events of moderate magnitude where surface deformation is hidden in Global Navigation Satellite System (GNSS) noise. However, in subduction zones where no clear relationship between tremor and slow slip occurrence is observed, these methods cannot be applied, and we need other methods to be able to better detect and quantify slow slip. Wavelets methods such as the Discrete Wavelet Transform (DWT) and the Maximal Overlap Discrete Wavelet Transform (MODWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted differences of a time series vary from one period to the next. In this paper, we use wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia. We use detrended GNSS data, apply the MODWT transform and stack the wavelet details over several nearby GNSS stations. As an independent check on the timing of slow slip events, we also compute the cumulative number of tremor in the vicinity of the GNSS stations, detrend this signal, and apply the MODWT transform. In both time series, we can then see simultaneous waveforms whose timing corresponds to the timing of slow slip events. We assume that there is a slow slip event whenever there is a positive peak followed by a negative peak in the wavelet signal. We verify that there is a good \textbf{agreement} between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data for northern Cascadia. The wavelet-based detection method effectively detects events of magnitude higher than 6 as determined by independent event catalogs (e.g. ~\citep{MIC_2019}). As a demonstration of using the wavelet analysis in a region without significant tremor, we also analyze GNSS data from New Zealand and detect slow slip events that are spatially and temporally close to those detected previously by other studies.

@@

Slow slip events are new phenomena discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks ~\citep{VER_2010,SCH_2010,JIA_2012,WAL_2012}. As with ordinary earthquakes, slow slip events represent slip on a fault, for instance the plate boundary between a tectonic plate subducting under another tectonic plate. However, they take a much longer time (several days to several years) to happen relative to ordinary earthquakes. They have a relatively short recurrence time (months to years) compared to the recurrence time of regular earthquakes (up to several hundreds of years), allowing scientists to observe and study many complete event cycles, which is typically not possible to explore with traditional earthquake catalogs ~\citep{BER_2011}. A slow slip event on the plate boundary is inferred to happen when there is a reversal of the direction of motion at GNSS stations, compared to the secular interseismic motion. Slow slip events have been observed in many places ~\citep{BER_2011,AUD_2016}, such as Cascadia ~\citep{BAR_2020}, Nankai ~\citep{NIS_2013}, Alaska ~\citep{LI_2016}, Costa Rica ~\citep{JIA_2012}, Mexico ~\citep{RAD_2012}, and New Zealand ~\citep{WAL_2020}. \\

In many places, tectonic tremor is also observed in relation to slow slip, but the spatial \textbf{agreement} between tremor and slow slip may vary along the strike of the plate boundary ~\citep{HAL_2018}. Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases. Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs) ~\citep{SHE_2007_nature}, which are small magnitude earthquakes (M $\sim$ 1) whose frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz). In subduction zones such as Nankai and Cascadia, tectonic tremor observations \textbf{agree} spatially and temporally with slow slip observations ~\citep{ROG_2003,OBA_2004}. Due to this \textbf{agreement}, these paired phenomena have been called Episodic Tremor and Slip (ETS). However, this is not always the case. For instance, in northern New Zealand, tremor is more challenging to detect, and seems to be located downdip of the slow slip on the plate boundary ~\citep{TOD_2016}. In Alaska, the tremor zone only partially overlaps the long-term slow slip zone and there does not appear to be any temporal \textbf{agreement} between tremor and slow slip occurrence ~\citep{WEC_2016}. \\

In Cascadia, there are robust signals in both slow slip and tremor ~\citep{HAW_2013}. This is also the case in Nankai ~\citep{HIR_2008}, where tiltmeters are used instead of GNSS. It is thus possible to use tremor as a proxy to observe slow slip events that are not directly observed in the GNSS data. For instance, ~\citet{AGU_2009} studied 23 ETS events in Cascadia with more than 50 hours of tectonic tremor. For all these events, they computed both the GPS-estimated moment release and the cumulative number of hours of tectonic tremor recorded. They observed a linear relationship between moment release and number of hours of tremor for slow slip events of moment magnitude 6.3 to 6.8. Based on this linear relationship, it is possible to infer the existence of smaller slow slip events of magnitude 5-6 occurring simultaneously with smaller tremor bursts of duration 1 to 50 hours occurring in between the big ETS events, and for which there is no detectable signal in the GPS data. \\

@@

Finally, ~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine relative vertical displacements associated with slow slip events between 1996 and 2011. Their main idea is that the tidal level measured at a given gauge is the sum of a noise component at multiple timescales (tides, ocean and atmospheric noise) and an uplift signal due to the slow slip events. The noise component is assumed to be coherent between all tidal gauges, while the tectonic uplift signal is different provided that the gauges are far enough from each other. By stacking the tidal records after removing tides, the uplift signals cancel each other while the noise signal is amplified. By stacking the components at different time scales of the DWT decomposition, instead of stacking the raw tidal record, each of the components of the noise at different time scales is retrieved and can then be removed from the raw records to obtain the uplift signal. Due to the relative location of the tidal gauges at Port Angeles and Port Townsend compared to the slow slip region on the plate boundary, a slow slip event should result in uplift in Port Angeles (western part) and in subsidence in Port Townsend (eastern part). Indeed, the authors were able to clearly see a difference in the sign of the uplift at these two tidal gauges. \\

In our study, we use a similar approach to previous studies with a different reasoning. We only stack signals at nearby GPS stations, assuming that the east-west displacement due to the slow slip events will then be the same at each of the GPS stations considered. We suppose that some of the noise component is different at each GPS station and will be eliminated by the stacking. Finally, we assume that the noise and the longitudinal displacement due to the slow slip events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the slow slip events. We use wavelet methods to analyze GPS and tremor recordings of slow slip events in Cascadia. Our objective is to verify that there is a good \textbf{agreement} between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data. We thus want to demonstrate that the wavelet-based detection method can be applied to detect slow slip events that may currently be obscured using standard methods. Finally, we apply the method to GNSS data in New Zealand and successfully detect several slow slip events without needing to rely on the tremor data. \\

@@

\textbf{The boundary conditions at the two edges of the time series will affect the wavelet coefficients. For the MODWT, if we denote $L$ the length of the base wavelet filter used for the wavelet decomposition (in our study, we used a Least Asymmetric wavelet filter of length $L = 8$, see ~\citep{PER_2000}, section 4.8, page 107), the length of the wavelet filter at level $j$ used to compute the wavelet detail $D_j$ is:}

\begin{equation*}
L_j = \left( 2^j - 1 \right) \left( L - 1 \right) + 1
\end{equation*}

\textbf{The wavelet coefficients of the detail al level $j$ affected by the boundary conditions at the edges would then be the coefficients with indices $t = 0 , \cdots , L_j - 2$ or $t = N - L_j + 1 , \cdots, N - 1$ (see ~\citep{PER_2000}, section 5.11, page 199). We get $L_j = 442$ for $j = 6$, $L_j = 890$ for $j = 7$ and $L_j = 1786$ for $j = 8$. In practice, the part of the wavelet details affected by the boundary conditions is much shorter than that. We compared the wavelet details computed when using only the data between 2008 and 2012 and the wavelet details computed when using the entire time series from 2000 to 2021 (Figure S1 in the Supplementary Material). Even at level 8 only about 6 months of data on each side are effected by the boundary conditions.}

@@

The DWT and MODWT methods must be used on a continuous time series, without gaps in the recordings. To deal with the gaps in the GNSS recordings, we simply replace the missing values by interpolation. The value for the first day for which data are missing is equal to the mean of the five days before the gap. The value for the last day for which data are missing is equal to the mean of the five days after the gap. The remaining missing values are computed by doing a linear interpolation of the first and the last values and adding a Gaussian noise component with mean zero and standard deviation equal to the standard deviation of the whole time series. We verify how the wavelet details may be affected by looking at a GPS time series without missing values and compared the wavelet details with and without removing some data points. Station PGC5 recorded continuous 1390 days between 2009 and 2013 without any missing values. We first computed the wavelet details without missing values. Then, we removed ten neighboring values, replaced them using the method described above (linear interpolation plus Gaussian noise), and computed the wavelet details with the replaced values. Figure S2 in the Supplementary Material shows a comparison of the two wavelet details for two different locations of the missing values. We can see that there are visible differences in the time series itself, and in the details at the smallest levels of the wavelet decomposition. However, the differences between the wavelet details with and without missing values get smaller and smaller with increasing levels of details, and are barely visible for the levels that are most relevant (levels 6 and above). We thus conclude that we can easily replace the missing values in the GNSS time series without introducing false detections of slow slip events. \\

We then applied the wavelet filtering to real GPS data. Figure 2 shows the longitudinal displacement for GPS station PGC5, located in southern Vancouver Island, the details of the wavelet decomposition for levels 1 to 8, and the smooth. In the data, we can see a sharp drop in displacement whenever there is a documented slow slip event. For levels 5 to 8, which correspond to time scales 16, 32, 64 and 128 days, we can see in the details a positive peak followed by a negative peak whenever there is a drop in displacement in the data. We thus verify that the wavelet method can detect steps in the time series associated with slow slip events. \\

To increase the signal-to-noise ratio and better detect slow slip events, we stack the signal from several neighboring GPS stations. We choose to focus on GPS stations located close enough to the tremor zone to get a sufficiently high amplitude of the slow slip signal. We choose 16 points along the 40 km depth contour of the plate boundary (model from ~\citet{PRE_2003}) with spacing equal 0.1 degree in latitude (red triangles on Figure 3). Then we took all the GPS stations located in a 50 km radius for a given point, compute the wavelet details for the longitudinal displacement of each station, and stack each detail over the GPS stations. We thus have a stacked detail for each level 1 to 10 of the wavelet decomposition. \\

To assess the success of the wavelet decomposition for detecting slow slip events in GPS time series, we validate the approach by comparing to an independent proxy for slow slip events. We took all the tremor epicenters located within a 50 km radius centered on one of the 16 locations marked by red triangles on Figure 3. Then we computed the cumulative number of tremor within this circle. Finally, we removed a linear trend from the cumulative tremor count, and applied the wavelet transform. Because of the preprocessing applied to the tremor data before that wavelet transform, \textbf{the measurement unit associated with the corresponding wavelet details is the fraction of tremor in a day divided by the total number of days. The average value is 1 divided by the total number of days.} Figure 4 shows an example of the wavelet decomposition for the third northernmost location on Figure 3 (which is closest to GPS station PGC5). Contrary to what happens for the GPS data, we see a sharp increase in the time series whenever there is a tremor episode, which translates into a negative peak followed by a positive peak in the wavelet details.

\section{Application to data from Cascadia}

We stacked the 8th level detail of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. The result is shown in the top panel of Figure 5, where each line represents one of the locations along strike. To better highlight the peaks in the wavelet details, we highlighted in red the time intervals where the amplitude of the stacked detail is higher than a threshold, and in blue the  time intervals where the amplitude of the stacked detail is lower than minus the threshold. To compare the GPS signal with the tremor signal, we plotted the 8th level detail of the wavelet decomposition of the tremor count on the bottom panel of Figure 5. We multiplied by -1 the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks. In the tremor catalog from the PNSN, there are 17 tremor events with more than 150 hours of tremor recorded. The events are summarized in Table 1. The time of the event is the start date plus half the duration of the event. \\

@@

We can then evaluate the quality of the detections obtained with our method by plotting a receiver operating characteristic curve (ROC curve). The ROC curve is widely use for binary classification problems in statistics and machine learning. We calculate an ROC value by varying the values of the threshold (here the two thresholds used to convert the GPS and the tremor time series into trinary time series), computing the corresponding values of the true positive rate and the false positive rate (equal to 1 minus the specificity), and plotting the true positive rate as a function of the false positive rate. If the classification was made randomly, all the points would fall on the first diagonal. If the classifier was perfect, the corresponding point would fall on the top left corner of the graph with true positive rate equal to 1 and false positive rate equal to 0. The bigger the area under the curve, the better the classification method is. \\

As the slow slip events are better seen on levels 6, 7 and 8 of the wavelet decomposition, we first add the wavelet details corresponding to levels 6 to 8, and transform the resulting time series into a trinary time series. We apply this transform to both the GPS and the tremor time series with varying thresholds. We then plot the ROC curve on Figure 8, each dot representing a different threshold. The corresponding sums of the wavelet details for the GPS data and the tremor data are shown on Figure 9. We can see that there is a trade-off between sensitivity and specificity as we vary the threshold. If we decrease the false positive rate, we also decrease the number of true events detected. If we increase the number of true events detected, we also increase the false positive rate. If we increase the threshold for the tremor, the curve goes farther away from the first diagonal, that is we get better classification results. If we increase the threshold for the GPS, the false positive rate and the the number of events detected decrease. In Figure 9, we have chosen thresholds for the GPS time series and the tremor time series such that the specificity is higher than 0.75 (that is the false positive rate is lower than 0.25), and the sensitivity is the highest possible, that is we have chosen the thresholds corresponding to the dot that is farthest from the diagonal, which is random. \\

In addition to the magnitude 6 events discussed above,  ~\citet{MIC_2019} have also identified several magnitude 5 events using a variational Bayesian Independent Component Analysis (vbICA) decomposition of the signal. As we expect smaller magnitude events to be more visible at smaller time scales of the wavelet decomposition (level 5), we verify for all these events whether a signal can be seen at the same time as the time given in their catalog. Most of these magnitude 5 events are also sub-events of bigger magnitude 6 events. Table 2 summarizes for each event its timing, its number and its magnitude as indicated in the catalog from ~\citet{MIC_2019}, and whether it is part of a bigger magnitude 6 event. Figure 10 shows the 5th level detail wavelet decomposition of the GPS data. Red lines show the timing of the big slow slip events from Table 1, and blue lines show the timing of the small slow slip events from Table 2. \\

All 14 events that are sub-events of a bigger event are visible at level 5. However, this may be because the bigger events are clearly seen at levels 6 to 8, and also at smaller time scales. The one small event that is not part of a bigger event (Winter 2009) is visible at level 5 of the wavelet decomposition. However, some other events that are not in the catalog of ~\citet{MIC_2019}'s catalog are also visible in late 2007, early 2010, early 2012, and early 2020. Therefore, it is difficult to differentiate between a robust detection and a false detection, and to conclude whether the method can indeed detect events of magnitude 5. \\

In Figure 9, we see four smaller events that are not in the catalog of ~\citet{MIC_2019}: at about 2007.5, there is a negative peak followed by a positive peak (that is an event in the opposite direction of what would be expected from slow slip), at about 2010.2, 2012.2 and 2020.2, there are positive peaks followed by negative peaks for all the sixteen locations studied in this paper. These events are highlighted in Figure S4 in the Supplementary Information. Looking back at the original GPS data, there is a small increase in the displacement in the eastern direction that lasts about one or two months at about 2007.5. However, the direction of the displacement does not correspond to a slow slip event, and another cause should be found to explain this signal. There is a decrease in displacement that lasts several months at about 2010.2. This transient may correspond to a long duration slow slip event. There is a small decrease in displacement at about 2012.2. Its amplitude is small but the duration and direction correspond to a slow slip event, so this transient could be a very small slow slip event. Finally, there is also a small decrease in displacement at about 2020.2 that is difficult to interpret. \\

Due to the short distances between the GPS stations and the locations of the red triangles on the map from Figure 3, the same station could be used multiple times for the stacking at different locations. When considering two different locations, the stacking is thus made over an overlapping number of stations. Table 3 summarizes the number of stations and the number of overlapping stations for each location on Figure 3. We hypothesize that the small displacement in the eastern direction seen at about 2007.5 could be due to a misbehaving station common to several locations. However, several GPS stations indeed show an increase in the displacement in the eastern direction at about 2007.5. There are many missing data around that time, so it is difficult to conclude. \\

Another possibility is that common mode signals could stack constructively across GNSS stations and produce peaks in the wavelet details that are actually due to non-tectonic signals. We computed common mode signals for different latitude bins (each bin has width equal to half-a-degree of latitude) following the same method as \citet{NUY_2021}. We first stacked all the time series for the stations in each latitude bin that are located more than 100 km east of the 40 km depth contour of the plate boundary. We assume that these stations are not sensitive to the deformation on the plate interface. We then apply a yearly moving average to each common mode signal in order to remove any leftover noise. The common mode signal was then removed from the GNSS time series depending on each site’s latitude. Figure S3 in the Supplementary Information shows the corresponding sum of the stacks of the 6th, 7th and 8th wavelet details obtained from the resulting time series. The common modes seem to have little impact on the results and do not explain the additional four small events that we noted in Figure 9. \\

In order to convert our filtered eastward displacement time series into a slow slip event catalog we note that red bars represent displacements exceeding a threshold of 0.8 mm (east), and blue marks displacements less than minus -0.8 mm (west).  During times with no slow slip GPS stations on the overriding plate are pushed slowly eastward by the locked subducting plate.  Slow slip events represent GPS motion \textbf{towards} the west.  Thus, we infer that slow slip events happen when red bars are immediately followed by blue bars in the wavelet details.  We have identified everywhere that this has happened and mark it with a green line in Figure 11 and as a row in Table 4.  We find 17 possible SSEs by this method using filtered GPS data only. For each of these 17 events we determine the time difference between the mid time of the GPS catalog and the nearest time from the tremor catalog (Table 1). These time differences are in column 6 (Table 4).  Every event in the GPS catalog has a match in the tremor catalog except for the tremor event at 2010.15. There is also only one event in the tremor catalog that is not in the GPS catalog. It occurs at 2014.65 with a duration of 15 days and 190 hours of tremor. It occurs 0.25 years after the nearest GPS event. There are also two marginal events in the tremor catalog with time differences of 0.13 and 0.10 years, but those are also among the smaller events with 162 and 193 hours of tremor.

\section{Application to data from New Zealand}

We now apply our wavelet-based method to detect slow slip events in New Zealand, a location where \textbf{the spatial and temporal agreement between tremor and slow slip is not as good as in} other subduction zones. The tectonics of the North Island of New Zealand are dominated by the westward subduction of the Pacific Plate under the Australian Plate at the Hikurangi Trench. Two types of slow slip events have been observed at the Hikurangi margin. Shallow (10-15 km depth), shorter (1-3 weeks), and usually smaller (Mw 6.3-6.8) slow slip events have been observed every 18-24 months in the northern part of the margin. Deeper (35-60 km depth), longer (12-18 months), and larger (Mw 7.0) slow slip events have been observed every 5 years in the southern part of the margin ~\citep{WAL_2010,TOD_2016}. The detection of tremor has been elusive in northern Hikurangi. ~\citet{DEL_2009} observed an increase in the rate of microseismicity downdip of the 2004 Gisborne slow slip event. More recently, however, ~\citep{KIM_2011} detected a low level of tremor activity that increased during the 2010 Gisborne slow slip event. As was the case for the microearthquakes, the source of the tremor was located downdip of the slow slip patch determined from GNSS data. ~\citep{IDE_2012} detected tremor downdip of the location of two deep slow slip events observed by ~\citet{WAL_2013} in 2006 and 2008. However, contrary to ETS events in Cascadia and Nankai, the tremor activity did not seem to increase during the slow slip events. ~\citet{TOD_2016}  detected tremor associated with most of the shallow slow slip events between 2010 and 2015, and located downdip of the geodetically inferred slip area. They also detected deeper tremor between 20 and 50 km depth with unclear origin. They hypothesized that these tremor may be related to undetected deep long-term slow slip events. \\
